\documentclass[../main.tex]{subfiles}
\begin{document}

\section{Simple Graphs}

\subsection{Definitions}

\textbf{Graphs} consist of vertices and edges. $G = (V, E)$

\textbf{Vertices} (sing. Vertex; also called Nodes) are the "dots". Vertices are represented by the set $V = \{a, b, c\}$.

\textbf{Edges} are the "lines". They are represented by the set of sets $E = \{\{a, b\}, \{b, c\}\}$

\textbf{Adjacent Vertices} are two vertices that are directly joined by an edge.

An edge is \textbf{incident} to the vertices is joins.

\textbf{Degree} of a vertex is the number of edges that are incident to it. Denoted by $deg(v)$.

\textbf{Complete Graph} has $n$ vertices and has an edge between every two vertices (total of $\frac{n(n-1)}{2}$). Denoted by $K_n$.

\textbf{Empty Graph} has at least one vertex and no edges.

\textbf{Line Graph} has $n$ vertices that are connected in a line by $n-1$ edges. Denoted by $L_n$.

\textbf{Cycle Graph} has $n$ vertices that are connected in a circle by $n$ edges. Denoted by $C_n$.

\textbf{Isomorphic Graphs} are two graphs that have the same number of vertices and edges and are arranged in the same way, so they "look the same". They may only be different in the names of vertices. (However, the same graph can often be drawn in lots of different ways.)

\textbf{Subgraphs} consist of a subset of the vertices and edges of some other graph.

\textbf{Weighted Graphs} are simple graphs that associate a real number, called weight, to each edge.

\textbf{Adjacency Matrix} represents a graph with $n$ vertices as a $n \times n$ matrix where the values are 1 if there is an edge between two nodes, and 0 otherwise. (For weighted graphs the weight is used instead of 1.)

\textbf{Bipartite Graphs} divide their vertices into two sets $L$ and $R$. Every edge is incident to a vertex in $L$ and $R$.

\textbf{Regular Graphs} have the same degree for all vertices.

\textbf{Acyclic Graphs} have no cycles.

\textbf{Trees} are connected acyclic graphs.

\textbf{Leaf} is a node of degree $1$ in a tree.

\textbf{Ordered Binary Trees} are trees where each node has at most two children labeled left-child and right-child.

\textbf{Spanning Tree} is a tree which is a subgraph of a connected graph $G$ that contains all the nodes of $G$ and a subset of its edges.

\textbf{Minimum (Weight) Spanning Tree (MST)} is a spanning tree of a weighted graph that has the smallest possible sum of edge weights.

\textbf{Walk} is a sequence of edges of a simple graph (graphs with multiple edges require a more strict definition) that are all connected. The same edge can occur multiple times in a single walk. Also called \textit{path}, in which case path is called \textit{simple path}.

\textbf{Path} (also \textit{simple path}) is a walk with no repeated nodes.

\textbf{Closed Walk} is a walk that starts and ends at the same node. Also called \textit{cycle} in which case cycle is called \textit{simple cycle}.

\textbf{Cycle} (also \textit{simple cycle}) is a closed walk with no repeated nodes (except the start and end nodes) and length at least $3$.

\textbf{Euler Walk} is a walk that traverses every edge exactly once.

\textbf{Euler Tour} is a closed walk that traverses every edge exactly once. In other words, it is an Euler walk that starts and ends at the same node.

\textbf{Hamiltonian Path} is a path that visits every node in a graph exactly once.

\textbf{Hamiltonian Cycle} is a cycle that visits every node in a graph exactly once.

\textbf{Connected Nodes} have a path between them. Every node is connected to itself by a path of length $0$.

\textbf{Connected Graph} has a path between every pair of nodes.

\textbf{Connected Component} is a subgraph of a graph and consists of some node and all nodes that are connected to it as well as the edges that connect them.

\textbf{k-Connected Graph} (also \textit{k-vertex connected graph}) is a graph that remains connected after removing any $k-1$ vertices.

\textbf{k-Edge Connected Graph} is a graph that remains connected after removing any $k-1$ edges.

\subsection{Matching Problems}

A \textit{matching} in a graph $G$ is a set of edges such that no two edges in the set share a vertex. A matching \textit{covers} a set of vertices $L$, iff every vertex in $L$ is incident to an edge of the matching. A matching is \textit{perfect} if every vertex is incident to an edge in the matching.

The set $N(S)$ called \textit{neighbors} of a set, $S$, of vertices contains all vertices adjecent to some vertex in $S$. $S$ is a \textit{bottleneck} if $|S| > |N(S)|$.

\textbf{Hall's Theorem:}
Let $G$ be a bipartite graph with vertex partition $L, R$. There is a matching in $G$ that covers $L$ iff no subset of $L$ is a bottleneck.

Every regular bipartite graph has a perfect matching.

\subsection{Coloring}

A \textit{coloring} or \textit{valid coloring} is an assignment of a color to each vertex such that no two adjecent vertices have the same color. A graph that has a coloring that uses at most $k$ colors is said to be \textit{k-colorable}. The minimum value of $k$ for which a graph $G$ has a valid $k$-coloring is called its \textit{chromatic number}, $\chi(G)$.

A graph with maximum degree at most $k$ is $(k+1)$-colorable. 

\subsection{Traveling Salesperson Problem}

The Traveling Salesperson Problem (TSP) is a problem in which a salesperson must visit a set of cities and return to the starting city. The goal is to find the shortest possible route that visits each city exactly once.

In other words, the TSP is a problem of finding a Hamiltonian cycle of minimum weight in a weighted graph.

\subsection{Trees}

Properties of trees:

\begin{enumerate}
    \item Any connected subgraph is a tree.
    \item There is a unique simple path between any two vertices.
    \item Adding an edge to a tree creates a cycle.
    \item Removing any edge disconnects the graph.
    \item If the tree has at least $2$ nodes, then it has at least $2$ leaves.
    \item The number of nodes in a tree is one more than the number of edges.
\end{enumerate}

\section{Directed Graphs}

\subsection{Communication Networks}

These are attributes used to describe communication networks:

\textbf{Network Diameter} is the distance between the two nodes that are farthest apart.

\textbf{Switch Size} is the number of inputs and outputs of the largest switch. For example, a $3 \times 3$ switch has $3$ inputs and $3$ outputs.

\textbf{Switch Count} is the number of switches in the network.

\textbf{Congestion} is the maximum number of paths that pass through a single switch in the worst case routing scenario.

The following table shows the attributes of some common communication networks:

\begin{center}
\begin{tabular}{l|c|c|c|c}
network              & diameter      & switch size  & switch count    & congestion \\
\hline
Complete Binary Tree & $2\log N + 2$ & $3 \times 3$ & $2N - 1$        & $N$        \\
\hline
2-D array            & $2N$          & $2 \times 2$ & $N^2$           & $2$        \\
\hline
butterfly            & $2 + \log N$  & $2 \times 2$ & $N(1 + \log N)$ & $\sqrt{N}$ or $\sqrt{N/2}$ \\
\hline
Benes                & $1 + 2\log N$ & $2 \times 2$ & $2N\log N$      & $1$        \\
\end{tabular}
\end{center}

\end{document}

