\documentclass[../main.tex]{subfiles}
\begin{document}

\section{Proofs}

Link to MIT OCW: \href{https://ocw.mit.edu/courses/6-042j-mathematics-for-computer-science-fall-2010/}{6.042 -- Mathematics for Computer Science (2010)}

\subsection{Terminology}

\begin{tabular}{ p{2.5cm} p{13cm} }
  \textbf{Proposition} & a statement that is either true or false \\
  \textbf{Predicate}  & a proposition whose truth depends on one or more
  variables e.g. "n is a perfect square" depends on the value of n \\
  \textbf{Axiom} & a proposition that is universially accepted as true, base
  for further proofs \\
  \textbf{Theorem} & an important true proposition \\
  \textbf{Lemma} & a preliminary proposition useful for proving later proposition \\
  \textbf{Corollary} & a proposition that follows in just a few logical steps
  from a theorem \\
\end{tabular}

\subsection{Notation}

\begin{tabular}{ p{2cm} p{8cm} p{4cm} }
  \textbf{Notation} & \textbf{Read as} & \textbf{Example} \\
  $\forall$ & "for all" & $\forall{n}\in\mathbb{N}.\ f(n)$ is prime. \\
  $\exists$ & "there exists" & $\exists{n}\in\mathbb{N}.\ n > 10.$ \\
  $\frac{\text{antecedents}}{\text{consequent}}$ & "if antecedents are proved, then the consequent (or conclusion) is also proved" & $\frac{P,\ P \implies Q}{Q}$ \\
\end{tabular}

\subsection{Propositions}

\subsubsection{Validity}

A proposition is \textit{valid} if it is always true (for any possible assignment of true and false to individual propositional variables).

Example:

$$
P \lor \neg P
$$

Both possible assignments result in the proposition being true, thus it is valid.

On the other hand:

$$
P \lor Q
$$

This proposition is false if both $P$ and $Q$ are false, thus it is not valid.

\subsubsection{Satisfiability}

A proposition is \textit{satisfiable} if there is an assignment of truth values to its individual propositional variables, such that the proposition is true.

Example:

$$
P \lor Q
$$

This proposition is true if either $P$ or $Q$ is true, thus it is satisfiable.

On the other hand:

$$
P \land \neg P
$$

There exists no assignment such that the proposition is true, thus it is not satisfiable.

\subsection{Logical Deductions (or Inference Rules)}

This inference rule is called \textit{modus ponens}:

$$
\frac{P,\ P \implies Q}{Q}
$$

There are also other inference rules:

$$
\frac{P \implies Q,\ \neg Q}{\neg P}
$$

$$
\frac
{P \implies Q,\ Q \implies R}
{P \implies R}
$$

$$
\frac{\neg P \implies \neg Q}{Q \implies P}
$$

\subsection{Patterns of Proof}

Proofs typically start with the word "Proof" and end with either "QED" or $\blacksquare$.

\subsubsection{Proving an Implication}
\label{sec:prove_implication}

\paragraph{Method 1: Assume P is true}

In order to prove that this is a pretty long text and it will take more than one line. jk, maybe now it will be long enough:

\begin{enumerate}
  \item Write, "Assume $P$."
  \item Show that $Q$ logically follows.
\end{enumerate}

\paragraph{Method 2: Prove the Contrapositive}
The implication $P \implies Q$ is equivalent to its \textit{contrapositive} $\neg Q \implies \neg P$.

\begin{enumerate}
  \item Write, "We prove the contrapositive:" and then state the contrapositive.
  \item Proceed as in method 1.
\end{enumerate}

\subsubsection{Proving an "If and Only If"}

\paragraph{Method 1}

$P \iff Q$ is equivalent to the two statements $P \implies Q$ and $Q \implies P$.

\begin{enumerate}
  \item Write, "We prove $P \implies Q$ and vice-versa."
  \item Write, "First, we show $P \implies Q$." Do this by one of the methods in Section~\ref{sec:prove_implication}.
  \item Write, "Now, we show $Q \implies P$." Again, do this by one of the methods in Section~\ref{sec:prove_implication}.
\end{enumerate}

\paragraph{Method 2}

\begin{enumerate}
  \item Write, "We construct a chain of if-and-only-if implications."
  \item Prove $P$ is equivalent to a second statement which is equivalent to a third statement and so on until you reach $Q$.
\end{enumerate}

\subsubsection{Proof by Cases}

A complicated proof can be divided into multiple cases in order to prove each case separately. It's important to also show that the proof covers all the cases.

\begin{enumerate}
  \item Write, "This proof is by case analysis."
  \item Prove each case individually (split further into subcases if necessary).
  \item Show that all the cases are covered and conclude that the initial proposition holds if all the cases hold.
\end{enumerate}

\subsubsection{Proof by Contradiction}

Also called \textit{indirect proof}. This approach is always viable, however, direct proofs are generally preferable when they are available.

In order to prove a proposition $P$ by contradiction:

\begin{enumerate}
  \item Write, "We use proof by contradiction."
  \item Write, "Suppose $P$ is false."
  \item Deduce something known to be false (a logical contradiction).
  \item Write, "This is a contradiction. Therefore, $P$ must be true."
\end{enumerate}

\subsection{Induction}

Proofs using the \textit{Well Ordering Principle}, \textit{Induction} and \textit{Strong Induction} are equivalent. This means, that a proof using any of these methods can be transformed into any of the other methods. One method might be easier or more natural than the other ones depending on the problem.

\subsubsection{Well Ordering Principle}

"Every \textit{nonempty} set of \textit{nonnegative integers} has a smallest element." However, this is not true for sets like $\emptyset$, $\mathbb{Z}^-$ and $\mathbb{R}$.

The following template can be used for a Well Ordering Proof:

\begin{enumerate}
  \item Define the set, $C$, of counterexamples to $P$ being true. $C ::= \{n \in \mathbb{N}\ |\ P(n)\ \text{is false}\}$.
  \item Use a proof by contradiction and assume that $P$ is nonempty.
  \item Reach a contradiction (somehow)---often by showing how to use $n$ to find another member of $C$ that is smaller than $n$.
  \item Conclude that $C$ must be empty, that is, no counterexamples exist.
\end{enumerate}

\subsubsection{Ordinary Induction}
\label{sec:ordinary_induction}

To use the Induction Principle for a proof, follow these steps:

\begin{enumerate}
  \item State that the proof uses induction.
  \item Define the predicate $P(n)$, also called the \textit{induction hypothesis}. It often follows directly from the proposition we want to prove. If the induction hypothesis includes multiple variables, we need to define which variable serves as $n$.
  \item Prove that $P(0)$ is true. This is called the \textit{base case}.
  \item Prove that $P(n) \implies P(n+1)$ for every nonnegative integer $n$. This is called the \textit{inductive step}. To prove this, assume $P(n)$ and show that $P(n+1)$ follows (see section \ref{sec:prove_implication}).
  \item Conclude that $P(n)$ is true for all nonnegative $n$.
\end{enumerate}

\subsubsection{Invariants}

The Induction Principle can be applied in the form of \textit{invariants}. An invariant is a property that is preserved for every possible state in a state machine. This is often used in computer science to prove that a program maintains a desired property.

In order to construct an invariant proof, follow these steps:

\begin{enumerate}
  \item Define $P(t)$ to be the predicate that some property \textsc{nice} holds after step $t$.
  \item Show that $P(0)$ is true, namely that \textsc{nice} holds for the start state.
  \item Show that $\forall t \in \mathbb{N}.\ P(t) \implies P(t+1)$, namely, that for any $t \ge 0$, if \textsc{nice} holds after step $t$, it must also hold after the following step.
\end{enumerate}

\subsubsection{Strong Induction}

The steps for using strong induction are mostly the same as the ones for ordinary induction listed in section \ref{sec:ordinary_induction}.

The only differences are:

\begin{itemize}
  \item It should be stated that the proof uses strong induction.
  \item $P(0),\ P(1),\ ...,\ P(n)$ can all be assumed to be true instead of only $P(n)$ during the inductive step.
\end{itemize}

\end{document}
