\documentclass[../main.tex]{subfiles}

\firstpageheader{6.042}{Problem Set 2}{Page \thepage\ of \numpages}
\runningheader{6.042}{Problem Set 2}{Page \thepage\ of \numpages}

\begin{document}
\begin{questions}

\question We will show that any sequence of five distinct integers will contain a \textit{3-chain} (subsequence of three integers, which is monotonically increasing or decreasing (not necessarily contiguous)).
\begin{parts}
  \part Assume that $a_1 < a_2$. Show that if there is not \textit{3-chain} in our sequence, then $a_3$ must be less than $a_1$. (Hint: consider $a_4$!)
  \begin{solution}

    There are 3 possibilities to insert $a_3$:

    Case 1: $a_1 < a_2 < a_3$. In this case we immediately have a 3-chain.

    Case 2: $a_1 < a_3 < a_2$. Now, $a_4$ can be either greater or lower than $a_3$. If $a_4 < a_3$ we have the 3-chain $a_2 > a_3 > a_4$. Otherwise, we have the 3-chain $a_1 < a_3 < a_4$.

    Case 3: $a_3 < a_1 < a_2$. This is the only case where it's possible to have no 3-chain up to $a_4$. In either $a_3 < a_4 < a_1 < a_2$ or $a_3 < a_1 < a_4 < a_2$ there is no 3-chain.
  \end{solution}

  \part Using the previous part, show that if $a_1 < a_2$ and there is no 3-chain in our sequence, then $a_3 < a_4 < a_2$.
  \begin{solution}

    If $a_4 < a_3$, there is a 3-chain: $a_4 < a_3 < a_1$, since $a_3 < a_1$ by part (a).
    
    Similarly, if $a_2 < a_4$, there is a 3-chain: $a_1 < a_2 < a_4$.

    Therefore, $a_3 < a_4 < a_2$ (two possible sequences shown in part (a)).
  \end{solution}

  \part Assuming that $a_1 < a_2$ and $a_3 < a_4 < a_2$, show that any value of $a_5$ must result in a 3-chain.
  \begin{solution}

    $a_5$ can be either greater or lower than $a_4$. If $a_5 < a_4$ there is a 3-chain $a_5 < a_4 < a_2$. Otherwise, there is also a 3-chain $a_3 < a_4 < a_5$. Therefore, any value of $a_5$ results in a 3-chain.
  \end{solution}

  \part Using the previous parts, prove by contradiction that any sequence of five distinct integers must contain a 3-chain.
  \begin{solution}

    Parts (a) through (c) show that assuming $a_1 < a_2$ there is no possible sequence that does not contain a 3-chain. By symmetry, the same holds for $a_2 < a_1$. Therefore, any sequence of five distinct integers must contain a 3-chain. $\blacksquare$
  \end{solution}
\end{parts}

\question Prove by either the Well Ordering Principle or Induction that for all nonnegative integers, $n$:

$$
\sum_{i=0}^{n} i^3 = \left(\frac{n(n+1)}{2}\right)^2
$$

\begin{solution}

  Proof by induction. Let the induction hypothesis $P(n)$ for $n >= 0$ be that the above statement holds for $n$.

  Base case: $P(0) = \left(\frac{0(0+1)}{2}\right)^2 = 0 = \sum_{i=0}^{0} i^3$. Thus, the base case holds.
  
  Inductive step: We assume $P(n)$ and show that $P(n+1)$ follows. We add $(n+1)^3$ to the right side of $P(n)$ and expand the terms:
  $$
  \left(\frac{n(n+1)}{2}\right)^2 + (n+1)^3
  $$
  $$
  = \frac{(n(n+1))^2 + 2^2(n+1)^3}{2^2}
  $$
  $$
  = \frac{n^2(n^2+2n+1) + 4(n^3+3n^2+3n+1)}{2^2}
  $$
  $$
  = \frac{n^4+2n^3+n^2+4n^3+12n^2+12n+4}{2^2}
  $$
  $$
  = \frac{n^4+6n^3+13n^2+12n+4}{2^2}
  $$

  Now, we expand the terms of $P(n+1)$:
  $$
  \left(\frac{(n+1)(n+2)}{2}\right)^2
  $$
  $$
  = \frac{(n+1)^2(n+2)^2}{2^2}
  $$
  $$
  = \frac{(n^2+2n+1)(n^2+4n+4)}{2^2}
  $$
  $$
  = \frac{n^4+4n^3+4n^2+2n^3+8n^2+8n+n^2+4n+4}{2^2}
  $$
  $$
  = \frac{n^4+6n^3+13n^2+12n+4}{2^2}
  $$

  Thus, we showed that:
  $$
  \sum_{i=0}^{n+1} i^3 = \left(\frac{n(n+1)}{2}\right)^2 + (n+1)^3 = \left(\frac{(n+1)(n+2)}{2}\right)^2
  $$
  Which means that $P(n) \implies P(n+1)$.

  We conclude by induction that $P(n)$ holds for all $n >= 0$. $\blacksquare$
\end{solution}

\question Prove the theorem: If fewer than $n$ students in class are initially infected, the whole class will never be completely infected.
\begin{solution}
  Proof by Induction. Let $P(k)$ for all $k >= 0$ be the induction hypothesis, that the perimeter of infected students after $k$ steps is at most $4n$.

  Base case: $P(0)$ is true, because initially we have less than $n$ students, which means the perimeter must be less than $4n$.

  Inductive step: We assume $P(k)$ to show that $P(k+1)$ follows. By the induction hypothesis, the perimeter of infected students after $k$ steps is less than $4n$. A new student can only be infected, if it is adjacent to two, three or four infected students. These cases change the perimeter in the following way:

  \begin{itemize}
    \item 2 adjacent: we remove 2 units and add 2 units = net change of 0
    \item 3 adjacent: we remove 3 units and add 1 unit = net change of -2
    \item 4 adjacent: we remove 4 units and add 0 units = net change of -4
  \end{itemize}
  This means that the perimeter can never increase, thus it is still less than $4n$ after $k+1$ steps.

  By induction, we can conclude that $P(k)$ holds for all $k >= 0$.

  Finally, since the perimeter of the whole class is $4n$, and the perimeter of the infected students can not increase, the whole class can never get infected if initially fewer than $n$ students are infected. $\blacksquare$
\end{solution}

\question Find the flaw in the bogus proof that $a^n=1$ for all nonnegative integers $n$, where $a$ is a nonzero real number.

\begin{solution}

  The flaw is in the indctive step $P(0) \implies P(1)$. In this step we have:

  $$
  a^{0+1}=\frac{a^0 \cdot a^0}{a^{0-1}} = \frac{1 \cdot 1}{1} = 1
  $$

  However, this incorrectly assumes $P(-1)$ while $P(n)$ only holds for $n >= 0$. Thus the proof is invalid.
\end{solution}

\question Prove that for all $n \in \mathbb{N}$, $G_n = k^n - 2^n$.

\begin{solution}

  Proof by strong induction. Let $P(k)$ for all $k \in \mathbb{N}$ be the proposition that $G_k = 3^k - 2^k$.

  Base case: $P(0)$ holds since $3^0 - 2^0 = 0 = G_0$. $P(1)$ holds since $3^1 - 2^1 = 1 = G_1$.

  Inductive step: We assume $P(k-2)$ and $P(k-1)$ for purposes of strong induction. We want to prove that:

  $$
  5G_{k-1} - 6G_{k-2} = 3^k - 2^k
  $$

  By $P(k-2)$ and $P(k-1)$ we can plug in:

  $$
  5(3^{k-1} - 2^{k-1}) - 6(3^{k-2} - 2^{k-2})
  $$
  $$
  = 5 \cdot 3^{k-1} - 5 \cdot 2^{k-1} - \frac{6}{3} \cdot 3^{k-1} + \frac{6}{2} \cdot 2^{k-1} 
  $$
  $$
  = 3 \cdot 3^{k-1} - 2 \cdot 2^{k-1}
  $$
  $$
  = 3^k - 2^k
  $$

  Thus, $P(k-2)$ and $P(k-1)$ imply $P(k)$ and we conclude by strong induction that $P(k)$ holds for every $k \in \mathbb{N}$. $\blacksquare$

\end{solution}

\end{questions}
\end{document}
